### 2021-04-19 input/experiments.tex
meow
small improvements everywhere

This section aims at highlighting the properties of \NAME that enables
adaptive scoped broadcast in dynamic networks. Most importantly, it
aims at answering the following questions: Does \NAME allows the
system to reach consistent partitioning? Does \NAME scales well in
large networks? Does \NAME improves scoped broadcast in
inter-autonomous systems? \TODO{Rework this.} We performed the
experimentation using PeerSim, a simulator written in Java that allows
researchers to evaluate and assess distributed algorithms in large
scale networks~\cite{montresor2009peersim}. The code is available on
the Github platform at
\url{https://github.com/Chat-Wane/peersim-partition}.\TODO{Anonoymous.}

>>>>

This section aims at highlighting the properties of \NAME that enables
adaptive scoped broadcast in dynamic networks. It answers the
following questions: What is the overhead of reaching consistent
partitioning using \NAME in large scale networks? Does \NAME is
relevant in Edge infrastructures where \processes are geo-distributed
into clusters? We performed the experimentation using PeerSim, a
simulator written in Java that allows researchers to evaluate and
assess distributed algorithms in large scale
networks~\cite{montresor2009peersim}. The code is available on the
Github platform at
\TODO{\url{https://github.com/Chat-Wane/peersim-partition}}.

### 2021-04-19 input/proposal.tex
small improvements everywhere

\begin{theorem}[\underline{B}est \underline{e}ventual \underline{f}orwarding $(BEF) \implies CP$]
  Assuming reliable communication links where a correct \process $q$
  eventually receives the message $m$ \underline{s}ent to it by a
  \process $p$ ($s_{pq}(m) \implies r_{q}(m)$); a monotonically
  growing system ($G^t \subseteq G^{t+1}$); \processes eventually
  reach consistent partitioning if each \process eventually forwards
  its best known partition.
\end{theorem}

@@

Algorithm~\ref{algo:add} shows the instructions that implement such
consistent partitioning when
\begin{inparaenum}[(i)]
\item weights are scalar values,
\item \processes only add new partitions to the system,
\item and \processes never crash nor leave the system.
\end{inparaenum}
Figure~\ref{fig:add} illustrates its behavior on a system comprising 4
\processes $a$, $b$, $c$, and $d$. \Process~$a$ and \Process~$d$
become the sources of their partition. They \NAMEB a notification
\underline{a}dd message: $\alpha_a^0$ and $\alpha_d^0$. They
initialize their own state with the lowest possible bound $0$ (see
Line~\ref{line:lowestbound}), and send a message to each of their
neighbors by accumulating the corresponding edge weight (see
Line~\ref{line:accumulator}). In Figure~\ref{fig:addC}, \Process~$b$
receives $\alpha_{d}^{1}$. Since it improves its own partition
distance, it keeps it and forwards it to its neighbors. In
Figure~\ref{fig:addD}, \Process~$b$ discards $\alpha_{a}^{2}$, for it
does not improve its partition distance. \Processes $c$ and $d$ will
never hear of Source~$a$. Ultimately, \processes discard last
transiting messages. Despite the obvious lack of traffic optimization,
the system reach consistent partitioning.

@@

At any time, a \process can become a source, hence adding a new
partition to the system. This partition eventually includes all
\processes that are closer from this new source than any other
else. \Processes naturally converge towards their respective best
partition by only piggybacking a monotonically increasing distance in
forwarded messages.

@@

\texttt{Del}ete operations add a new notion of order between events,
and most importantly between message deliveries. Delete operations
implicitly state that all preceding events become obsolete, and that
all messages originating from these preceding events convey stale
control information.

@@

\begin{definition}[Last win order]
  When a \process $p$ broadcasts two messages $b_p(m) \rightarrow
  b_p(m')$, no \process $q$ can deliver $m$ if it has delivered $m'$:
  $\not\exists q \in V$ with $d_q(m') \rightarrow d_q(m)$, for $m'$
  convey \emph{stale} control information.
\end{definition}

@@

Unfortunately, stale control information about deleted partitions may
impair the propagation of both
\begin{inparaenum}[(i)]
\item notifications about actual sources, and
\item notifications about other deleted partitions.
\end{inparaenum}
Figure~\ref{fig:problem} highlights such consistency issue with
dynamic partitions, even in contexts where \processes have FIFO links,
\ie where \processes receive the messages in the order of their
sending. Both \Process~$a$ and \Process~$d$ add then delete their
partition concurrently. \Process~$c$ receives, delivers, and forwards
$\alpha_d^3$ followed by $\delta_d$. In the meantime, \Process~$b$
receives, delivers, and forwards $\alpha_a^2$. In
Figure~\ref{fig:problemC}, both \Process~$c$ and \Process~$d$ deliver
the message about Partition $P_a$, for they did not have any known
partition at receipt time. On the contrary, \Process~$b$ delivers
$\alpha_d^1$, for it improves its distance to a known
source. \Process~$b$ then blocks \Process~$a$'s removal notification
$\delta_a$. It never reaches \Process~$c$ nor \Process~$d$. Also,
\Process~$c$ and \Process~$d$ do not deliver any $\alpha_d$ or
$\delta_d$ since they already delivered it. The system converges to an
inconsistent state where some \processes assume they belong to a
partition while it does not exist anymore. Naive propagation of
$\alpha$ and $\delta$ messages is insufficient to guarantee consistent
partitioning when each \process can add or delete its partition at any
time. \Processes need to exploit their local knowledge gathered over
message receipts.

@@

One could guarantee consistent partitioning by always propagating the
$\delta$ messages corresponding to the $\alpha$ messages it propagated
before. In Figure~\ref{fig:problem}, it means that as soon as
\Process~$b$ forwards $\alpha_a^2$, it assumes that its neighbors
\Process~$c$ and \Process~$d$ may need the notification of removal
$\delta_a$ if it exists. However, such solution also implies that
\processes generate traffic not only related to their current
partition, but also related to partitions they belong to in the
past. Instead, we propose to use the delivery order at each \process
to detect possible inconsistencies and solve them. Together,
\processes eventually remove all stale control information (transiting
messages and local states) of the system leaving room for propagation
of messages about up-to-date partitions.

@@

\begin{proof}
  \Processes deliver and forward the best sum of weights they
  received. The only case where a \process would block such epidemic
  propagation would be when it received and delivered a message from a
  partition that has since then been deleted but with a better sum of
  weights than newly received ones. Removing \emph{forever} all such
  stale messages about deleted partitions would allow \processes to
  propagate their best up-to-date partition again, eventually reaching
  consistent partitioning together.
\end{proof}

>>>>

\begin{theorem}[\label{theo:bef}\underline{B}est \underline{e}ventual
    \underline{f}orwarding $(BEF) \implies CP$]
%
Assuming reliable communication links where a correct \process $q$
eventually receives the message $m$ \underline{s}ent to it by a
\process $p$ ($s_{pq}(m) \implies r_{q}(m)$); a monotonically growing
system ($G^t \subseteq G^{t+1}$); \processes eventually reach
consistent partitioning if each \process eventually forwards its best
known partition.
\end{theorem}

@@

Algorithm~\ref{algo:add} shows the instructions that implement such
consistent partitioning when
\begin{inparaenum}[(i)]
\item weights are scalar values,
\item \processes only add new partitions to the system,
\item and \processes never crash nor leave the system.
\end{inparaenum}
Figure~\ref{fig:add} illustrates its behavior on a system comprising 4
\processes $a$, $b$, $c$, and $d$. \Process~$a$ and \Process~$d$
become the sources of their partition. They \NAMEB a notification
\underline{a}dd message: $\alpha_a^0$ and $\alpha_d^0$. They
initialize their own state with the lowest possible bound $0$ (see
Line~\ref{line:lowestbound}), and send a message to each of their
neighbors by accumulating the corresponding edge weight (see
Line~\ref{line:accumulator}). In Figure~\ref{fig:addC}, \Process~$b$
receives $\alpha_{d}^{1}$. Since it improves its own partition
distance, it keeps it and forwards it to its neighbors. In
Figure~\ref{fig:addD}, \Process~$b$ discards $\alpha_{a}^{2}$, for it
does not improve its partition distance. \Processes $c$ and $d$ will
never acknowledge that Source~$a$ exists. Ultimately, \processes
discard last transiting messages. Despite the obvious lack of traffic
optimization, the system reach consistent partitioning.

@@

%% At any time, a \process can become a source, hence adding a new
%% partition to the system. This partition eventually includes all
%% \processes that are closer from this new source than any other
%% else. \Processes naturally converge towards their respective best
%% partition by only piggybacking a monotonically increasing distance in
%% forwarded messages.

@@

Delete operations add a new notion of order between events, and most
importantly between message deliveries. Delete operations implicitly
state that all preceding events become obsolete, and that all messages
originating from these preceding events convey stale control
information.

@@

\begin{definition}[\label{def:lwo}Last win order]
  When a \process $p$ broadcasts two messages $b_p(m) \rightarrow
  b_p(m')$, no \process $q$ can deliver $m$ if it has delivered $m'$:
  $\not\exists q \in V$ with $d_q(m') \rightarrow d_q(m)$, for $m'$
  convey \emph{stale} control information.
\end{definition}

@@

Unfortunately, stale control information as stated in
Definition~\ref{def:lwo} may impair the propagation of both
\begin{inparaenum}[(i)]
\item notifications about actual sources, and
\item notifications about deleted partitions.
\end{inparaenum}
Figure~\ref{fig:problem} highlights such consistency issue with
dynamic partitions, even in contexts where \processes have FIFO links,
\ie where \processes receive the messages in the order of their
sending. Both \Process~$a$ and \Process~$d$ add then delete their
partition concurrently. \Process~$c$ receives, delivers, and forwards
$\alpha_d^3$ followed by $\delta_d$. In the meantime, \Process~$b$
receives, delivers, and forwards $\alpha_a^2$. In
Figure~\ref{fig:problemC}, both \Process~$c$ and \Process~$d$ deliver
the message about Partition $P_a$, for they did not have any known
partition at receipt time. On the contrary, \Process~$b$ delivers
$\alpha_d^1$, for it improves its distance to a known
source. \Process~$b$ then blocks \Process~$a$'s removal notification
$\delta_a$. It never reaches \Process~$c$ nor \Process~$d$. Also,
\Process~$c$ and \Process~$d$ do not deliver any $\alpha_d$ or
$\delta_d$ since they already delivered it. The system converges to an
inconsistent state where some \processes assume they belong to a
partition while it does not exist anymore. Naive propagation of
$\alpha$ and $\delta$ messages is insufficient to guarantee consistent
partitioning when each \process can add or delete its partition at any
time. \Processes need to exploit their local knowledge gathered over
message receipts.

@@

One could guarantee consistent partitioning by always propagating the
$\delta$ messages corresponding to the $\alpha$ messages it propagated
before. In Figure~\ref{fig:problem}, it means that as soon as
\Process~$b$ forwards $\alpha_a^2$, it assumes that its neighbors
\Process~$c$ and \Process~$d$ may need the notification of removal
$\delta_a$ if it exists. However, such solution also implies that
\processes generate traffic not only related to their current
partition, but also related to partitions they belong to in the
past. This would prove overly expensive in dynamic systems where
\processes join or leave the system, create or delete partitions, at
any time. Instead, we propose to use the delivery order at each
\process to detect possible inconsistencies and solve them. Together,
\processes eventually remove all stale control information (transiting
messages and local states) of the system leaving room for propagation
of messages about up-to-date partitions.

@@

\begin{proof}
  \Processes deliver and forward the best sum of weights they
  received. Figure~\ref{fig:proofA} depicts the only case where a
  \process would block such epidemic propagation: the \process
  received and delivered a message from a partition that has since
  then been deleted but with a better sum of weights than newly
  received ones. Removing \emph{forever} all such stale messages about
  deleted partitions would allow \processes to propagate their best
  up-to-date partition again, eventually reaching consistent
  partitioning together, as stated by Theorem~\ref{theo:bef}.
\end{proof}
